---
title: "Intro to R-workshop"
author: "Stephanie Long and Morgan Middlebrooks"
date: "4/3/2020"
output:
  pdf_document: default
  html_document: default
---

# Section 1: Programming in R Intro
# Authored by Stephanie Long


R in its most basic form can be used as a calculator.

```{r}
# addition
5 + 3

# subtraction
5 - 3 

# multiplication
5 * 3 

# division
5 / 3

# find the remainder
5 %% 3 

# exponentiation
3^2

# square root
sqrt(25) 
```


You can use R to assign values to variables to store for further use

```{r}
# assign 5 to x
x = 5

x

# another way to assign 4 to y
y <- 4

y
# yet another way

10 -> z

z

# assign 6 to new variable called "x4"
x4 = 6

# now we can use variables we created above to make a calculation
x * x4

#now we assign values to two variables
test_object <- 1.5
test_object_2 <- 2

# we can add variables together
test_object_2 + test_object
```

You can also test logical operators

```{r}
# using our stored values we can compare variables
# is z greater than x?
z > x  

# is x less than y?
x < y 

# is z greater than or equal to x?
z >= x 

# is z equal to y?
z == y

# is the value of z - 6 equal to y?
z - 6 == y  

# is z not equal to y?
z != y 
```


Finally, variables (or objects) in R are stored as certain data types
```{r}
# 1) numeric
typeof(x)

# 2) character
l <- "this is a list"
typeof(l)

# 3) logical
typeof(TRUE)

# 4) list (contains multiple data types)
my_list <-list("a" = 2.5, "b" = TRUE, "c" = 1:3)
typeof(my_list)

# these are all vectors (1 dimension objects)

# there are also 2 dimensional objects

# 1 matrix (contains rows and columns)
my_matrix <- matrix(data = c(1, 2, 3,
                             4, 5, 6,
                             7, 8, 9), nrow = 3, ncol = 3)
my_matrix
str(my_matrix)

# 2 data frame (contains observations (rows) and variables (columns))
my_dataframe <- as.data.frame(my_matrix)
str(my_dataframe)

# what sets this apart from matrices is that you can call on variables by their name using "$"
my_dataframe$V1

```


# Section 2: ANOVA Basics
# Authored by Morgan Middlebrooks


Understand what an ANOVA is even doing/ what sort of data is appropriate for it. 

It's important to understand what TYPE of data you need in ANOVA. An ANOVA is basically just seeing if there is statistically significant difference between the groups of a IV on some outcome variable. 

So you need the IV to be categorical and the DV to be continuous. 

There are certain packages that will make this easier.

```{r}
library(tidyverse)

# first we will load in our data
load(file = "Movie Data.Rdata")

#shows us what's inside the data
head(movie_data)

# need to remove incomplete cases of data aka "NA"
movie_data <- movie_data[complete.cases(movie_data$genre),]
```

Next we will subset the data.

We're going to pick the genres with the largest cell sizes that are closest to equal.

```{r}

## It looks like Adventure, Comedy, and Drama are our closest to equal cell
## sizes
group_by(movie_data, genre) %>%
  summarise(
    count = n())


anova.df <- subset(movie_data, genre %in% c('Adventure','Comedy', 'Drama'))

group_by(anova.df, genre) %>%
  summarise(
    count = n())

boxplot(total_gross ~ genre, data = anova.df, main = "Gross Revenue",
        ylab = "Total Gross Revenue", xlab = "Genre", col = "salmon")

outlier.invest <- ggplot(anova.df, aes(genre, total_gross)) + 
  stat_boxplot(geom = 'errorbar') +
  geom_boxplot(fill = "steelblue", colour = "black", outlier.colour = "red", outlier.shape = 4) +
  ggtitle("Outlier Investigation") + 
  theme_classic() +
  theme(plot.title = element_text(lineheight=.8, face="bold", hjust = .5, size = 12, colour = "black"))

outlier.invest

```

It looks like we have some pretty serious outliers. 

Now we will demonstrate one way of removing outliers

```{r}

outliers <- boxplot(anova.df$total_gross, plot=FALSE)$out

print(outliers)

anova.df.out <- anova.df[-which(anova.df$total_gross %in% outliers),]

outlier.invest.2 <- ggplot(anova.df.out, aes(genre, total_gross)) + 
  stat_boxplot(geom = 'errorbar') +
  geom_boxplot(fill = "steelblue", colour = "black", outlier.colour = "red", outlier.shape = 4) +
  ggtitle("Outlier Investigation") + 
  theme_classic() +
  theme(plot.title = element_text(lineheight=.8, face="bold", hjust = .5, size = 12, colour = "black"))

outlier.invest.2
```


We still have some after a basic outlier removal that look like outliers, but were not classified this way. There are lots of fancy ways you can check if those are TRUE outliers (using Cook's d for example), but we're not doing that today.

Notice how the bars are stretched more similarly now.

Now we will check the groups for equal variance since this is another important aspect of an ANOVA.
```{r}

car::leveneTest(total_gross ~ genre, data = anova.df.out)
```

There is a difference, but barely. A better scientist would not let this slide. The groups should have equal variance!

Now we can move on to the ANOVA and interpretation.

```{r}

mod <- aov(total_gross ~ genre, data = anova.df.out)

summary(mod)
```

There is a difference in genre....but WHERE YOU ASK?

Now we can do post-hoc tests to find where the significant difference in genres is.

```{r}

TukeyHSD(mod)

## Difference between comedy + adventure
## Difference between drama + adventure
## No difference in drama + comedy

group_by(anova.df, genre) %>%
  summarise(
    count = n(),
    mean = mean(total_gross, na.rm = TRUE),
    sd = sd(total_gross, na.rm = TRUE)
  )

## adventure makes way more money here
```

Check out boxplots below to really send message home that you're going to make more money on adventure.

```{r}
ggplot(anova.df.out, aes(genre, total_gross)) + 
  stat_boxplot(geom = 'errorbar') +
  geom_boxplot(fill = "steelblue", colour = "black", outlier.colour = "red", outlier.shape = 4) +
  ggtitle("Movie Monies") + 
  theme_classic() +
  theme(plot.title = element_text(lineheight=.8, face="bold", hjust = .5, size = 12, colour = "red"))
```

Looking at this plot you can see that Adventure's mean is higher than both comedy and drama, but comedy and drama aren't that different.

# Section 3: Data Visualization of Movie Genre Revenue
# Authored by Stephanie Long

We will continue to use the movie_data dataset from above but we need to clean up a few things

```{r}

# removing genre groups with less than 10 movies
large_sample_genre <- filter(movie_data, genre != "Black Comedy" & genre != "Concert/Performance" & genre != "Horror" & genre != "Western")
summary(large_sample_genre$genre)

## adding two columns to divide by 1 billion, so the y-axis of the plots makes more sense
large_sample_genre$total_gross_billion <- large_sample_genre$total_gross/1000000000
large_sample_genre$adj_gross_billion <- large_sample_genre$inflation_adjusted_gross/1000000000
```

Our first plot will be a simple boxplot

```{r}

plot1 <- ggplot(data = large_sample_genre) + # call our data layer to the plot
  geom_boxplot(aes(x = genre, # adding our boxplot data, this line adds our x-axis
                   y = total_gross_billion, # add our y-axis
                   fill = genre)) + # fill the boxes with color based on or genre groups
  theme(axis.text.x  = element_text(angle=75, vjust=0.5, size=9), # adjust the x-axis text so that it's angled and larger
        legend.position = "none", # remove the legend since you can see what groups the colors belong to
        plot.title = element_text(hjust = 0.5)) + # adjust our title to be in the center
  labs(x = "Genre Category", # re-title the x-axis label
       y = "Total Gross Revenue (Billions)", # re-title the y-axis label
       title = "Movie Genre Total Gross Revenue") # add the title
plot1
```

Our second plot will demonstrate two bar graphs including errorbars

The plot will show mean gross revenue for both adjusted and non-adjusted for inflation for each genre and include standard error for revenue type across genres as errorbars. Standard error is a measurement of how much a sample mean deviates from the population mean and is pretty standard for plotting errorbars.

We'll start with a cleaned up and organized dataset created from the movie_data

```{r}

load("Plot 2 cleaned data.Rdata")
str(plot_data2)

# same general set-up
plot2 <- ggplot(data = plot_data2, 
                aes(x = genre,
                    y = MRevenue, 
                    fill = Revenue_Type)) + # our colors here will reflect adjusted or total revenue
  geom_bar(stat = "identity") + # adding our bar graph by assigning the height of the bar to be our MRevenue
  theme_bw() + # adding a new more plain theme as our background
  theme(axis.text.x  = element_text(angle=75, vjust=0.5, size=9),
        legend.position = "none") +
  facet_wrap(~Revenue_Type) + # adding a "facet" so that there are two graphs plotted side by side
  labs(x = "Genre Category", 
       y = "Mean Gross Revenue (Billions)") +
  geom_errorbar(aes(ymin=MRevenue-SE, ymax=MRevenue+SE), width=.2, position=position_dodge(.9)) #adding our errorbars as standard error
plot2
```

Finally our third plot will compare the top 25 grossing movies comparing the adjusted and non-adjusted revenues.

```{r}

library("patchwork") # package for combining plots

#pulling just the top 25 total grossing movies
top_total_25 <- movie_data %>% 
  arrange(desc(total_gross)) %>% #arranges the data in descending order by "total_gross"
  slice(1:25) # slices just the top 25

top_total_25

#pulling just the top 25 grossing movies adjusted for inflation
top_adj_25 <- movie_data %>% 
  arrange(desc(inflation_adjusted_gross)) %>%
  slice(1:25)

top_adj_25

#combining the data and converting for graphing (don't worry about this part)
top_25 <- merge(top_total_25[,c(1,5)], top_adj_25[,c(1,6)], all = T, by = "movie_title")
top_25_pivot <- pivot_longer(top_25, cols = c("total_gross", "inflation_adjusted_gross"), names_to = "Type", values_to = "Revenue")

# we're going to make two separate graphs of our top 25 movies and then combined them

# plot 3A: top 25 based on "total_gross"
total <- ggplot(data = top_total_25, 
                aes(x = reorder(movie_title, total_gross), # we can arrange our bars to be descending from greatest to least
                    y = (total_gross/1000000))) + # we can divide the y axis so it's showing Millions instead of Billions
  geom_bar(stat = "identity", position = "dodge", fill = "turquoise2") + # change the color of our bars with "fill"
  theme_bw() + 
  coord_flip() + # we can flip our x and y coordinates
  theme(axis.title.y = element_blank(), # remove the y axis title since we all the movies are labeled
        legend.position = "none",
        plot.title = element_text(hjust = 0.5))+
  labs(y = "Total Revenue (Millions)", # make sure our labels go with our data
       title = "Top 25 Movies Based on Total Revenue")
total

# plot 3B: top 25 based on "inflation_adjusted_gross"
adj <- ggplot(data = top_adj_25, 
              aes(x = reorder(movie_title, inflation_adjusted_gross), y = (inflation_adjusted_gross/1000000000))) +
  geom_bar(stat = "identity", position = "dodge", fill = "salmon") +
  theme_bw() +
  coord_flip() +
  theme(axis.title.y = element_blank(), 
        legend.position = "none",
        plot.title = element_text(hjust = 0.5))+
  labs(y = "Total Revenue (Billions)",
       title = "Top 25 Movies Based on Inflation Adjusted Revenue")
adj

# combining plot 3A and 3B
plot3 <- total | adj
plot3
