---
title: "R-workshop"
author: "Stephanie Long"
date: "4/3/2020"
output: html_document
---

# Set working directory and load data

```{r}

load(file = "Movie Data.Rdata")
head(movie_data)

```


# 2: Understand what an ANOVA is even doing/ what sort of data is appropriate for it. 

## It's important to understand what TYPE of data you need in ANOVA. An ANOVA is
## basically just seeing if there is statistically significant difference
## between the groups of a IV on some outcome variable. 

## So you need the IV to be categorical and the DV to be continuous.

# Look at the structure of the data and the type of variables in it

```{r}
str(movie_data)

```

```{r}
summary(movie_data)

# looks like we havve NAs (no value) in the genre category which we are interested in

movie_data <- movie_data[complete.cases(movie_data$genre),]
summary(movie_data)

```

```{r}

# we should only look at genres with adequate samples
xtabs(~genre, data = movie_data)

#Black Comedy, Concert/Performance, Horror, and Western have less than 10 cases...

movie_data <- subset()

```

# What is a package?
# What is a boxplot?

```{r}

library("tidyverse")

ggplot(data = movie_data) +
  geom_boxplot(aes(x = genre, y = total_gross)) +
  theme(axis.text.x  = element_text(angle=75, vjust=0.7, size=9))

```
